---
title: "cleaning_and_setup"
author: "Errol Kaylor"
date: "`r Sys.Date()`"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(rio)
library(here)
library(janitor)
library(pdftools)
library(tabulapdf)
library(shinydashboard)
library(igraph)
library(readxl)
library(CurricularAnalytics)
library(visNetwork)
library(shiny)
```


```{r data import}
all_courses<- import(here("data/UG_course_prereq_split - UG_course_prereq.csv.csv")) %>% 
  clean_names() %>% 
  rename(course_id = "v1",
         has_prereq = "v2",
         course_dept = "v3",
         course_num = "v4") %>% 
  select(-c(5,6)) %>% 
  filter(course_id != "Common_Course_ID") %>% 
  filter(course_dept == "CS" | course_dept == "MATH")


prereqs <- import(here("data/Prereq_Coreq_AY2324 Cleaned.xlsx")) %>% 
  clean_names()
```


```{r visualization and cleaning}
  
sample_data <- read_excel("data/Untitled spreadsheet (1).xlsx") %>% 
  add_row(requisites = "CS299",Course = "CS299")

test_set<- prereqs %>% 
  filter(active_course_ind == "Y" & !grepl('Law',academic_period_desc) ) %>% 
  filter(subject=="CS" | subject == "MATH") %>% 
  group_by(course_identification) %>% 
  reframe(count_offered = n(),
          prereqtext = requisites_text) %>% 
  distinct()


all_courses <- full_join(all_courses,test_set, by=join_by(course_id==course_identification))


```

Need to create a parser for the requisitess
Aggregate Course Data
- Number of requisitess
- Frequency Offered
- Number of layers of requisitess
- Number of dependencies


-corequisite

Problems -- singletons, prereqs from outside departments!
```{r prereq parser}


#goal is generate a list of edges and the requirements for each edge to be active. 

all_courses_filtered<- all_courses %>% 
  select(-c(2,3,4,5)) %>% 
  na.omit()
all_courses$prereqtext[[1]]
all_courses$prereqtext[[2]]
layer_1 <- str_split(all_courses_filtered$prereqtext[[6]],pattern = " AND ")
layer_2<- str_split(flatten(layer_1),pattern = " OR ")




prereq_sample <- str_split(flatten(layer_2),pattern = " FOR LEVEL UG WITH MIN. GRADE OF | FOR LEVEL GR WITH MIN. GRADE OF | FOR MIN. SCORE OF") 

prereqs <- list_transpose(flatten(str_split(layer_1,pattern = " FOR LEVEL UG WITH MIN. GRADE OF | FOR LEVEL GR WITH MIN. GRADE OF | FOR MIN. SCORE OF")))  


prereq_transposed <- list_transpose(prereq_sample)

edge_list <- tibble(prereq = prereq_transposed[[1]],
                    grade_req = prereq_transposed[[2]],
                    course = all_courses$course_id[[2]])


edge_list$prereq<- str_remove(edge_list$prereq,"\\(|\\)")
edge_list$grade_req <- str_remove(edge_list$grade_req,"\\(|\\)")
```

```{r}
get_prereq_edges <-  function(all_courses,index){
  layer_0 <- str_split(all_courses$prereqtext[[index]],pattern = " AND ")
  layer_1 <- str_split(flatten(layer_0),pattern = " OR ")
  prereqs <- list_transpose(str_split(flatten(layer_1),pattern = " FOR LEVEL UG WITH MIN. GRADE OF | FOR LEVEL GR WITH MIN. GRADE OF | FOR MIN. SCORE OF"))

  edge_list <- tibble(prereq = prereqs[[1]],
                      course = all_courses$course_id[[index]],
                      grade_req = prereqs[[2]],
                      nums = str_extract(course,pattern = "\\d\\d\\d" ),
                      x = str_sub(nums,end=1),
         y = str_sub(nums,2)
                      )
  
  
  edge_list$prereq<- str_remove(edge_list$prereq,"\\(|\\)")
  edge_list$grade_req <- str_remove(edge_list$grade_req,"\\(|\\)")
  
  
  
  
  edge_list
}


all_courses_filtered<- all_courses_filtered  %>% 
  mutate(nums = str_extract(all_courses_filtered$course_id,pattern = "\\d\\d\\d" ),
         x = str_sub(nums,end=1),
         y = str_sub(nums,2))

edge_list <- tibble()
for (i in seq_along(all_courses_filtered[[1]])){
  print(i)
  edge_list <-rbind(get_prereq_edges(all_courses_filtered,i),edge_list)
}


```


Visualization Ideas
Overview/Course Analytics Definitions

Department/Degree View
-All courses in department
-

```{r}

#C <- curriculum_graph_from_csv(here("data/Untitled spreadsheet - Sheet1.csv"))

#term is used for horizontal spacing -- use department as chunk area (+20, then index)
#
#
generate_coords <- function(curriculum_graph) {
  coords <- matrix(ncol = 2)

  #track num courses in each dept/term, increment by 1, starting from 0
  #
  num_in_section <- array(0,dim = c(5,10))
  y_coord <- 0
  old_dept <- "not!"
  old_term <- 0
  for (i in curriculum_graph$node_list$id) {
    term <- as.numeric(curriculum_graph$node_list$term[i])
    label <- curriculum_graph$node_list$label[i]
    nums <- as.numeric(str_extract(label,pattern = "\\d\\d\\d"))
    dept <- str_extract(label,pattern ="[A-Z]+")
    y_starter <- case_when(
      dept == "CS" ~ 1,
      dept == "MATH" ~ 2,
      dept == "PHYS" ~ 3,
      dept == "PEO" ~ 4
    )

    section_depth <- num_in_section[y_starter,term]
    
    num_in_section[y_starter,term]<-section_depth+2
    
    ycoord <- (y_starter * 20) + section_depth
    
    
    coords <- rbind(coords, c(term,((y_starter * 20) + section_depth)))

  }

  coords <- stats::na.omit(coords)
  return(coords)
}

```




```{r shiny_dash}



ui <- dashboardPage(
  dashboardHeader(title = "UO Curriculum Analytics"),
  dashboardSidebar(
    sidebarMenu(
      menuItem("Dashboard", tabName = "dashboard", icon = icon("dashboard")),
      menuItem("Widgets", tabName = "widgets", icon = icon("th"))
    )
  ),
  dashboardBody(
    tabItems(
      # First tab content
      tabItem(tabName = "dashboard",
        fluidRow(
          box(visNetworkOutput("mynetworkid"))
          )
        )
      ),

      # Second tab content
      tabItem(tabName = "widgets",
        h2("Widgets tab content")
      )
    )
  )


server <- function(input, output) {

  output$mynetworkid <- renderVisNetwork({
    coords <- generate_coords(C)
visNetwork(C$node_list,C$edge_list,width="100%",
           submain = paste(
      "Total Structural Complexity:",
      C$sc_total,
      "Total Blocking Factor:",
      C$bf_total,
      "Total Delay Factor:",
      C$df_total
    )) %>% 
  visNodes(shape = "circle", 
           color = list(background = "lightgreen", 
                        border = "darkgreen",
                        highlight = "yellow")) %>%
  visEdges(arrows = "to",color = "green") %>% 
  visIgraphLayout(layout= "layout.norm",layoutMatrix = C_coords) %>% 
  visOptions(selectedBy = "group") %>% 
  visInteraction(navigationButtons = TRUE) %>% 
  visEvents(
      selectNode = "function(properties) {
      alert(' sc: ' + this.body.data.nodes.get(properties.nodes[0]).sc + ' cf: ' + this.body.data.nodes.get(properties.nodes[0]).cf + ' bf: ' + this.body.data.nodes.get(properties.nodes[0]).bf + ' df: ' + this.body.data.nodes.get(properties.nodes[0]).df);}"
    )
  })
}

shinyApp(ui, server)



```

#write my vignettes on what we're trying to do, have a comparison of two majors!

CS Overview

term Meaning/Definitions

Degree/Dept Comparisons




```{r}

raw_list <- read.csv(here("data/Untitled spreadsheet - Sheet1.csv"))
# Create list for storing pathways
Gc <- list()

set.seed(87460945)
for (i in 1:50) {
  # Generate degree pathway
  pathway <-
    data.frame(
      id = NA,
      label = NA,
      term = NA,
      requisites = NA
    )

  # First Year
  pathway <-
    rbind(pathway, cbind(
      subset(raw_list, label == "CS210")
    ))
  samp <- sample(c("MATH251", "MATH246"), 1)
  pathway <-
    rbind(pathway, cbind(subset(raw_list, label == samp)))
    samp <- sample(c("MATH252", "MATH247"), 1)
  pathway <-
    rbind(pathway, cbind(subset(raw_list, label == samp)))

  samp <- sample(c(TRUE, FALSE), 1)

    pathway <-
      rbind(pathway, cbind(
        subset(raw_list, label == "CS211")
      ))

  # samp <- c("PHYS 111", "PHYS 121")
  # pathway <-
  #   rbind(pathway, cbind(subset(ubco, label %in% samp), data.frame("term" =
  #                                                                          c(1, 2))))
  # 
  # pathway <-
  #   rbind(pathway, cbind(subset(
  #     ubco, label %in% c("COSC 111", "COSC 121")
  #   ), data.frame("term" = c(1, 2))))


  # Second Year
  pathway <-
    rbind(pathway, cbind(subset(
      raw_list,
      label %in% c("CS212",
                         "MATH231",
                         "CS314",
                         "MATH232",
                         "CS322")
    )))


  # Third and Fourth Year
  pathway <-
    rbind(pathway, cbind(subset(
      raw_list,
      label %in% c("CS313",
                         "CS315",
                         "CS330",
                         "CS415",
                         "CS425",
                         "CS422")
    )))

  #Math Requirements
  samp <- sample(c("MATH253","MATH341","MATH343","MATH425","MATH345"),1)
  pathway<-rbind(pathway,cbind(subset(raw_list,label == samp)))


  upper_year_data <-
    c("CS413","CS420","CS423","CS429","CS431","CS432","CS433","CS434","CS436","CS437","CS441","CS443","CS445","CS451","CS453","CS461","CS471","CS472","CS473")

  course <- c()
  while (length(unique(course)) < 5) {
      course <- c(course, sample(upper_year_data, 1))
  }

  pathway <-
    rbind(pathway, cbind(
      subset(raw_list, label %in% unique(course))
     ))

  pathway <- na.omit(pathway)
  rownames(pathway) <- 1:nrow(pathway)

  # Construct node and edge list

node_list <-
  data.frame(
    id = rownames(pathway),
    label = pathway$label,
    term = pathway$term
  )

edge_list <- data.frame(from = NA, to = NA)

for (node_id in node_list$id) {
    node_reqs <- subset(raw_list, id == node_id)$requisites
    req_ids <- ""
     if (node_reqs != "") {
       req_ids <- as.numeric(strsplit(node_reqs, ";")[[1]])
       for (req_id in req_ids) {
         if (!req_id %in% node_list$id){
           req_node <- subset(raw_list,id == req_id)
           node_list <- rbind(node_list,c(req_node$id,req_node$label,req_node$term))
         }
       }
#      node_list <- rbind(node_list,cbind(subset(raw_list, id %in% req_ids)))

      from <- rownames(subset(raw_list, id == node_id))
      to <- rownames(subset(raw_list, id %in% req_ids))
      if (length(to) > 1) {
        for (id in to) {
          edge_list <- rbind(edge_list, data.frame(from = id, to = from))
        }
      } else {
        edge_list <- rbind(edge_list, data.frame(from = to[1], to = from))
      }
     }
  }



# for (node in node_list$label) {
#   # node = "COSC 329"
#   str <- subset(pathway, label == node)$requisites
#   
#   
#   
#   
#   
# 
#   from <- rownames(subset(pathway, label == node))
#   to <- rownames(subset(pathway, label %in% str))
#   if (length(to) > 1) {
#     for (id in to) {
#       edge_list <- rbind(edge_list, data.frame(from = id, to = from))
#     }
#   } else {
#     edge_list <- rbind(edge_list, data.frame(from = to[1], to = from))
#   }
# 
# }
edge_list <- na.omit(edge_list)
node_list$id <- as.numeric(node_list$id)

node_list$new_id <- 1:32
new_id_list <- data.frame(
  id = NA
)
new_edge_list <- data.frame(from = NA, to = NA)
for (i in 1:40){
  old_val <- edge_list$from[i]

  new_val <- subset(node_list, id == old_val)$new_id
  new_id_list <- rbind(new_id_list,new_val)
}
new_id_list <- na.omit(new_id_list)

new_node_list <- data.frame(id = node_list$new_id,
                            label = node_list$label,
                            term = node_list$term)

new_edge_list <- data.frame(from = new_id_list,
                            to = edge_list$to) %>% 
  rename(from = id) %>% 
  mutate(from = as.character(from))

C <- curriculum_graph_from_list(new_node_list, new_edge_list)
Gc <- c(Gc, list(C))
}
```


```{r}
C<- curriculum_graph_from_list(new_node_list,new_edge_list)

C$node_list$group <- C$node_list$label %>% 
  map(str_extract,pattern = "[A-Z]+")

C_coords <- generate_coords(C)

```


```{r}
C_orig <- curriculum_graph_from_csv(here("data/Untitled spreadsheet - Sheet1.csv"))

C_orig$node_list$group <- C_orig$node_list$label %>% 
  map(str_extract,pattern = "[A-Z]+")

orig_coords <- generate_coords(C_orig)
```

