---
title: "cleaning_and_setup"
author: "Errol Kaylor"
date: "`r Sys.Date()`"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(rio)
library(here)
library(janitor)
library(pdftools)
library(tabulapdf)
library(shinydashboard)
library(igraph)
library(readxl)
library(CurricularAnalytics)
library(visNetwork)
library(shiny)
```


```{r data import}
all_courses<- import(here("data/UG_course_prereq_split - UG_course_prereq.csv.csv")) %>% 
  clean_names() %>% 
  rename(course_id = "v1",
         has_prereq = "v2",
         course_dept = "v3",
         course_num = "v4") %>% 
  select(-c(5,6)) %>% 
  filter(course_id != "Common_Course_ID") %>% 
  filter(course_dept == "CS" | course_dept == "MATH")


prereqs <- import(here("data/Prereq_Coreq_AY2324 Cleaned.xlsx")) %>% 
  clean_names()
```


```{r visualization and cleaning}
  
sample_data <- read_excel("data/Untitled spreadsheet (1).xlsx") %>% 
  add_row(Prerequisite = "CS299",Course = "CS299")

test_set<- prereqs %>% 
  filter(active_course_ind == "Y" & !grepl('Law',academic_period_desc) ) %>% 
  filter(subject=="CS" | subject == "MATH") %>% 
  group_by(course_identification) %>% 
  reframe(count_offered = n(),
          prereqtext = prerequisite_text) %>% 
  distinct()


all_courses <- full_join(all_courses,test_set, by=join_by(course_id==course_identification))


```

Need to create a parser for the prerequisites
Aggregate Course Data
- Number of prerequisites
- Frequency Offered
- Number of layers of prerequisites
- Number of dependencies


-corequisite

Problems -- singletons, prereqs from outside departments!
```{r prereq parser}


#goal is generate a list of edges and the requirements for each edge to be active. 

all_courses_filtered<- all_courses %>% 
  select(-c(2,3,4,5)) %>% 
  na.omit()
all_courses$prereqtext[[1]]
all_courses$prereqtext[[2]]
layer_1 <- str_split(all_courses_filtered$prereqtext[[6]],pattern = " AND ")
layer_2<- str_split(flatten(layer_1),pattern = " OR ")




prereq_sample <- str_split(flatten(layer_2),pattern = " FOR LEVEL UG WITH MIN. GRADE OF | FOR LEVEL GR WITH MIN. GRADE OF | FOR MIN. SCORE OF") 

prereqs <- list_transpose(flatten(str_split(layer_1,pattern = " FOR LEVEL UG WITH MIN. GRADE OF | FOR LEVEL GR WITH MIN. GRADE OF | FOR MIN. SCORE OF")))  


prereq_transposed <- list_transpose(prereq_sample)

edge_list <- tibble(prereq = prereq_transposed[[1]],
                    grade_req = prereq_transposed[[2]],
                    course = all_courses$course_id[[2]])


edge_list$prereq<- str_remove(edge_list$prereq,"\\(|\\)")
edge_list$grade_req <- str_remove(edge_list$grade_req,"\\(|\\)")
```

```{r}
get_prereq_edges <-  function(all_courses,index){
  layer_0 <- str_split(all_courses$prereqtext[[index]],pattern = " AND ")
  layer_1 <- str_split(flatten(layer_0),pattern = " OR ")
  prereqs <- list_transpose(str_split(flatten(layer_1),pattern = " FOR LEVEL UG WITH MIN. GRADE OF | FOR LEVEL GR WITH MIN. GRADE OF | FOR MIN. SCORE OF"))

  edge_list <- tibble(prereq = prereqs[[1]],
                      course = all_courses$course_id[[index]],
                      grade_req = prereqs[[2]],
                      nums = str_extract(course,pattern = "\\d\\d\\d" ),
                      x = str_sub(nums,end=1),
         y = str_sub(nums,2)
                      )
  
  
  edge_list$prereq<- str_remove(edge_list$prereq,"\\(|\\)")
  edge_list$grade_req <- str_remove(edge_list$grade_req,"\\(|\\)")
  
  
  
  
  edge_list
}


all_courses_filtered<- all_courses_filtered  %>% 
  mutate(nums = str_extract(all_courses_filtered$course_id,pattern = "\\d\\d\\d" ),
         x = str_sub(nums,end=1),
         y = str_sub(nums,2))

edge_list <- tibble()
for (i in seq_along(all_courses_filtered[[1]])){
  print(i)
  edge_list <-rbind(get_prereq_edges(all_courses_filtered,i),edge_list)
}


```


Visualization Ideas
Overview/Course Analytics Definitions

Department/Degree View
-All courses in department
-


```{r}
generate_coords <- function(curriculum_graph) {
  coords <- matrix(ncol = 2)

  old_term <- 1
  idx <- -1
  for (term in curriculum_graph$node_list$term) {
    if (old_term != term) {
      idx <- 0
      old_term <- term
    } else{
      idx <- idx + 1
    }
    coords <- rbind(coords, c(term, idx))
  }

  coords <- stats::na.omit(coords)
  return(coords)
}
```



```{r shiny_dash}



C <- curriculum_graph_from_csv(here("data/Untitled spreadsheet - Sheet1.csv"))

ui <- dashboardPage(
  dashboardHeader(title = "UO Curriculum Analytics"),
  dashboardSidebar(
    sidebarMenu(
      menuItem("Dashboard", tabName = "dashboard", icon = icon("dashboard")),
      menuItem("Widgets", tabName = "widgets", icon = icon("th"))
    )
  ),
  dashboardBody(
    tabItems(
      # First tab content
      tabItem(tabName = "dashboard",
        fluidRow(
          box(visNetworkOutput("mynetworkid", height = 250))
          )
        )
      ),

      # Second tab content
      tabItem(tabName = "widgets",
        h2("Widgets tab content")
      )
    )
  )


server <- function(input, output) {

  output$mynetworkid <- renderVisNetwork({
    coords <- generate_coords(C)
visNetwork(C$node_list,C$edge_list,width="100%") %>% 
  visEdges(arrows = "to") %>% 
  visIgraphLayout(layout= "layout.norm",layoutMatrix = coords)
  })
}

shinyApp(ui, server)

generate_coords(C)
```

#write my vignettes on what we're trying to do, have a comparison of two majors!



