---
title: "UO Curricular Analytics Shiny Dashboard"
author: "Errol Kaylor"
date: "`r Sys.Date()`"
output: html_document
runtime: shiny
---

```{r setup, include=FALSE}
library(tidyverse)
library(rio)
library(here)
library(janitor)
library(shinydashboard)
library(igraph)
library(readxl)
library(CurricularAnalytics)
library(visNetwork)
library(shiny)
library(kableExtra)
```


First part here is creating different degree pathways, drawing courses from a central list with the following columns:
id: Integer values for each course, starting from 1 and increasing in order.
label:Short Name for the course, e.g. "CS 201"
term: Integer representing which term a course should be taken in. 1 represents first term of their first year, 3 is the third term of their first year, 5 is the second term of their second year, and so on.
requisites: list of course ID's that lead to the course, e.g. 5;6;11;13 
```{r pathway creation}

#This dataset contains 130 courses, spread out across the Math, Physics, and Computer Science departments. 
#An important next step is ingesting this from the course catalog directly, rather than hand-coding the requisites.
#Link to Computer Science Major Requirements -- https://catalog.uoregon.edu/arts-sciences/school-computer-data-sciences/computer-science/ug-computer-science/#requirementstext 


raw_list <- read.csv(here("data/Untitled spreadsheet - Sheet1.csv"))
# Create list for storing pathways
Gpathway <- list()
Gc <- list()
set.seed(6402025) #Setting a seed for reproducibility.
for (i in 1:200) {
  # Generate degree pathway
  pathway <-
    data.frame(
      id = NA,
      label = NA,
      term = NA,
      requisites = NA
      #After adding more columns to the input dataset, add them here as well
    )
  #
  # First Year
  pathway <-
    rbind(pathway, cbind(
      subset(raw_list, label == "CS210")
    ))
  samp <- sample(c("MATH251", "MATH246"), 1)
  pathway <-
    rbind(pathway, cbind(subset(raw_list, label == samp)))
    samp <- sample(c("MATH252", "MATH247"), 1)
  pathway <-
    rbind(pathway, cbind(subset(raw_list, label == samp)))

  samp <- sample(c(TRUE, FALSE), 1)

    pathway <-
      rbind(pathway, cbind(
        subset(raw_list, label == "CS211")
      ))


  # Second Year
  pathway <-
    rbind(pathway, cbind(subset(
      raw_list,
      label %in% c("CS212",
                         "MATH231",
                         "CS314",
                         "MATH232",
                         "CS322")
    )))


  # Third and Fourth Year
  pathway <-
    rbind(pathway, cbind(subset(
      raw_list,
      label %in% c("CS313",
                         "CS315",
                         "CS330",
                         "CS415",
                         "CS425",
                         "CS422")
    )))

  #Math Requirements
  samp <- sample(c("MATH253","MATH341","MATH343","MATH425","MATH345"),1)
  pathway<-rbind(pathway,cbind(subset(raw_list,label == samp)))

  #Next section selects a subset of upper level courses to add to the pathway. 
  upper_year_data <-
    c("CS413","CS420","CS423","CS429","CS431","CS432","CS433","CS434","CS436","CS437","CS441","CS443","CS445","CS451","CS453","CS461","CS471","CS472","CS473")

  course <- c()
  while (length(unique(course)) < 5) {
      course <- c(course, sample(upper_year_data, 1))
  }

  pathway <-
    rbind(pathway, cbind(
      subset(raw_list, label %in% unique(course))
     ))

  pathway <- na.omit(pathway)
  rownames(pathway) <- 1:nrow(pathway)
  # Construct node and edge list

node_list <-
  data.frame(
    id = pathway$id,
    label = pathway$label,
    term = pathway$term
  )

edge_list <- data.frame(from = NA, to = NA)
#the following loop adds all requisites to the node list, and all associated edges.
#the loop only goes back one level (could rewrite to be more efficient) so we run it 3 times in case a there were multiple levels of requisites to sign up for a chosen course above.
for (node_id in node_list$id) {
    node_reqs <- subset(raw_list, id == node_id)$requisites
    req_ids <- ""
     if (node_reqs != "") {
       req_ids <- as.numeric(strsplit(node_reqs, ";")[[1]])
       for (req_id in req_ids) {
         if (!req_id %in% node_list$id){
          
           req_node <- subset(raw_list,id == req_id)
           node_list <- rbind(node_list,c(req_node$id,req_node$label,req_node$term))

         }
       }
#      node_list <- rbind(node_list,cbind(subset(raw_list, id %in% req_ids)))

      from <- rownames(subset(raw_list, id == node_id))
      to <- rownames(subset(raw_list, id %in% req_ids))
      if (length(to) > 1) {
        for (id in to) {
          edge_list <- rbind(edge_list, data.frame(from = id, to = from))
        }
      } else {
        edge_list <- rbind(edge_list, data.frame(from = to[1], to = from))
      }
     }
}


for (node_id in node_list$id) {
    node_reqs <- subset(raw_list, id == node_id)$requisites
    req_ids <- ""
     if (node_reqs != "") {
       req_ids <- as.numeric(strsplit(node_reqs, ";")[[1]])
       for (req_id in req_ids) {
         if (!req_id %in% node_list$id){
          
           req_node <- subset(raw_list,id == req_id)
           node_list <- rbind(node_list,c(req_node$id,req_node$label,req_node$term))

         }
       }
#      node_list <- rbind(node_list,cbind(subset(raw_list, id %in% req_ids)))

      from <- rownames(subset(raw_list, id == node_id))
      to <- rownames(subset(raw_list, id %in% req_ids))
      if (length(to) > 1) {
        for (id in to) {
          edge_list <- rbind(edge_list, data.frame(from = id, to = from))
        }
      } else {
        edge_list <- rbind(edge_list, data.frame(from = to[1], to = from))
      }
     }
}

for (node_id in node_list$id) {
    node_reqs <- subset(raw_list, id == node_id)$requisites
    req_ids <- ""
     if (node_reqs != "") {
       req_ids <- as.numeric(strsplit(node_reqs, ";")[[1]])
       for (req_id in req_ids) {
         if (!req_id %in% node_list$id){
          
           req_node <- subset(raw_list,id == req_id)
           node_list <- rbind(node_list,c(req_node$id,req_node$label,req_node$term))

         }
       }
#      node_list <- rbind(node_list,cbind(subset(raw_list, id %in% req_ids)))

      from <- rownames(subset(raw_list, id == node_id))
      to <- rownames(subset(raw_list, id %in% req_ids))
      if (length(to) > 1) {
        for (id in to) {
          edge_list <- rbind(edge_list, data.frame(from = id, to = from))
        }
      } else {
        edge_list <- rbind(edge_list, data.frame(from = to[1], to = from))
      }
     }
}




node_list$id <- as.numeric(node_list$id)
node_list$new_id <- seq_along(node_list$id)
new_id_list <- data.frame(
  id = NA
)
new_to_list <- data.frame(id= NA)

edge_list <- na.omit(edge_list)

new_edge_list <- data.frame(from = NA, to = NA)
new_node_list <- data.frame(id = NA, label = NA, term = NA)


#the id's from the original course list are out of order, here we link those old id's to new ones to create graphable node and edge lists.
for (i in 1:length(edge_list$from)){
  old_val <- edge_list$from[i]
  old_to_val <- edge_list$to[i]
  new_to_val <- subset(node_list,id == old_to_val)$new_id
  new_val <- subset(node_list, id == old_val)$new_id
  new_id_list <- rbind(new_id_list,new_val)
  new_to_list <- rbind(new_to_list,new_to_val)

}

new_id_list <- na.omit(new_id_list)
new_to_list <- na.omit(new_to_list)

new_node_list <- data.frame(id = node_list$new_id,
                            label = node_list$label,
                            term = node_list$term)



new_edge_list <- data.frame(from = new_id_list,
                            to = new_to_list$id) %>% 
  rename(from = id) %>% 
  mutate(from = as.character(from)) %>% 
  unique()

C <- curriculum_graph_from_list(new_node_list, new_edge_list)
Gc <- c(Gc, list(C))
Gpathway<- c(Gpathway, list(pathway))
}
```

```{r max and min paths}
# # Initialize variables for maximum and minimum values
max_total <- -Inf
min_total <- Inf
max_index <- NULL
min_index <- NULL
#
# # Iterate through Gc to find graphs in GC with maximum and minimal structural complexity.
for (i in seq_along(Gc)) {
  total <- sum(Gc[[i]]$sc_total)
  if (total > max_total) {
    max_total <- total
    max_index <- i
  }
  if (total < min_total) {
    min_total <- total
    min_index <- i
  }
}

C_max <- Gc[[max_index]]


C_min <- Gc[[min_index]]

C<-curriculum_graph_from_csv(here("data/Untitled spreadsheet - Sheet1.csv"))

#Add a column to the curriculum graphs showing the department each course belongs to.
C$node_list$group <- C$node_list$label %>% 
  map(str_extract,pattern = "[A-Z]+")

C_max$node_list$group <- C_max$node_list$label %>% 
  map(str_extract,pattern = "[A-Z]+")

C_min$node_list$group <- C_min$node_list$label %>% 
  map(str_extract,pattern = "[A-Z]+")
```



```{r functions}
#coordinates for the network visualization is a matrix with two columns and one row for each node in the network.
#coordinates are set to split each department into a different segment vertically, and horizontally by term. As courses are added to the same department and term, the y_coordinate is incremented.

generate_coords <- function(curriculum_graph) {
  coords <- matrix(ncol = 2)


  num_in_section <- array(0,dim = c(5,10))
  y_coord <- 0
  old_dept <- "not!"
  old_term <- 0
  for (i in curriculum_graph$node_list$id) {
    term <- as.numeric(curriculum_graph$node_list$term[i])
    label <- curriculum_graph$node_list$label[i]
    nums <- as.numeric(str_extract(label,pattern = "\\d\\d\\d"))
    dept <- str_extract(label,pattern ="[A-Z]+")
    y_starter <- case_when(
      dept == "CS" ~ 1,
      dept == "MATH" ~ 2,
      dept == "PHYS" ~ 3,
      dept == "PEO" ~ 4
    )

    section_depth <- num_in_section[y_starter,term]
    
    num_in_section[y_starter,term]<-section_depth+2
    
    ycoord <- (y_starter * 20) + section_depth
    
    
    coords <- rbind(coords, c(term,((y_starter * 25) + section_depth)))

  }

  coords <- stats::na.omit(coords)
  return(coords)
}

#used for printing information from meaningful courses in a given network. 
print_top_two_rows <- function(df, column) {
  ordered_df <- df[rev(order(df[[column]])), ]
  top_two <- head(ordered_df, 3)
  print(top_two)
}

```


```{r tables}
columns <- colnames(C_max$node_list[,c("bf","df","cf","sc")])
for (column in columns) {
  print(paste("Ordering by column:", column))
  print_top_two_rows(C_max$node_list, column)
}

columns <- colnames(C$node_list[,c("bf","df","cf","sc")])
for (column in columns) {
  print(paste("Ordering by column:", column))
  print_top_two_rows(C$node_list, column)
}

idx <- !(C_max$node_list$label %in% C_min$node_list$label)
kable(data.frame(Courses=C_max$node_list$label[idx]))


idx <- !(C_min$node_list$label %in% C_max$node_list$label)
kable(data.frame(Courses=C_min$node_list$label[idx]))
```

```{r shiny dash}
ui <- dashboardPage(
  dashboardHeader(title = "UO Curriculum Analytics"),
  dashboardSidebar(
    sidebarMenu(
      menuItem("Degree Path Analysis", tabName = "dashboard", icon = icon("dashboard")),
      menuItem("Overview", tabName = "overview", icon = icon("th"))
    )
  ),
  dashboardBody(
    tabItems(
      # First tab content
      tabItem(tabName = "dashboard",
        fluidRow(
          box(title ="Max CS Degree Path",visNetworkOutput("maxnetworkid")),
          box(title = "Min CS Degree Path",visNetworkOutput("minnetworkid"))
          ),
        fluidRow(
          box(title = "Max Path Courses Overview",tableOutput("max_only_courses")),
          box(title = "Min Path Courses Overview",tableOutput("min_only_courses"))
        )
        ),
      tabItem(tabName = "overview",
              fluidRow(
                box(title = "Curriculum Overview",visNetworkOutput("overallcnetworkid"))
              ))
      )
      )
    )


server <- function(input, output) {
  library(dplyr)
  library(kableExtra)
#Below generates the network visualization for a given curriculum graph, and surfaces important structural elements of the network.  
  output$maxnetworkid <- renderVisNetwork({
    max_coords <- generate_coords(C_max)
visNetwork(C_max$node_list,C_max$edge_list,width="100%",
           submain = paste(
      "Total Structural Complexity:",
      C_max$sc_total,
      "Total Blocking Factor:",
      C_max$bf_total,
      "Total Delay Factor:",
      C_max$df_total
    )) %>% 
  visNodes(shape = "circle", 
           color = list(background = "lightgreen", 
                        border = "darkgreen",
                        highlight = "yellow")) %>%
  visEdges(arrows = "to",color = "green") %>% 
  visIgraphLayout(layout= "layout.norm",layoutMatrix = max_coords) %>% 
  visOptions(selectedBy = "group") %>% 
  visEvents(
      selectNode = "function(properties) {
      alert('Course Code: ' + this.body.data.nodes.get(properties.nodes[0]).label + '; Structural Complexity: ' + this.body.data.nodes.get(properties.nodes[0]).sc + '; Centrality Factor: ' + this.body.data.nodes.get(properties.nodes[0]).cf + '; Blocking Factor: ' + this.body.data.nodes.get(properties.nodes[0]).bf + '; Delay Factor: ' + this.body.data.nodes.get(properties.nodes[0]).df);}"
    )
  })
  
  output$minnetworkid <- renderVisNetwork({
    min_coords <- generate_coords(C_min)
visNetwork(C_min$node_list,C_min$edge_list,width="100%",
           submain = paste(
      "Total Structural Complexity:",
      C_min$sc_total,
      "Total Blocking Factor:",
      C_min$bf_total,
      "Total Delay Factor:",
      C_min$df_total
    )) %>% 
  visNodes(shape = "circle", 
           color = list(background = "lightgreen", 
                        border = "darkgreen",
                        highlight = "yellow")) %>%
  visEdges(arrows = "to",color = "green") %>% 
  visIgraphLayout(layout= "layout.norm",layoutMatrix = min_coords) %>% 
  visOptions(selectedBy = "group") %>% 
  visEvents(
      selectNode = "function(properties) {
      alert('Course Code: ' + this.body.data.nodes.get(properties.nodes[0]).label + '; Structural Complexity: ' + this.body.data.nodes.get(properties.nodes[0]).sc + '; Centrality Factor: ' + this.body.data.nodes.get(properties.nodes[0]).cf + '; Blocking Factor: ' + this.body.data.nodes.get(properties.nodes[0]).bf + '; Delay Factor: ' + this.body.data.nodes.get(properties.nodes[0]).df);}"
    )
  })
  
  output$overallcnetworkid<- renderVisNetwork({
    coords <- generate_coords(C)
visNetwork(C$node_list,C$edge_list,width="100%",
           submain = paste(
      "Total Structural Complexity:",
      C$sc_total,
      "Total Blocking Factor:",
      C$bf_total,
      "Total Delay Factor:",
      C$df_total
    )) %>% 
  visNodes(shape = "circle", 
           color = list(background = "lightgreen", 
                        border = "darkgreen",
                        highlight = "yellow")) %>%
  visEdges(arrows = "to",color = "green") %>% 
  visIgraphLayout(layout= "layout.norm",layoutMatrix = coords) %>% 
  visOptions(selectedBy = "group") %>% 
  visEvents(
      selectNode = "function(properties) {
      alert('Course Code: ' + this.body.data.nodes.get(properties.nodes[0]).label + '; Structural Complexity: ' + this.body.data.nodes.get(properties.nodes[0]).sc + '; Centrality Factor: ' + this.body.data.nodes.get(properties.nodes[0]).cf + '; Blocking Factor: ' + this.body.data.nodes.get(properties.nodes[0]).bf + '; Delay Factor: ' + this.body.data.nodes.get(properties.nodes[0]).df);}"
    )
  })
  
  idx <- !(C_max$node_list$label %in% C_min$node_list$label)
  max_tab <- data.frame(Max_Only_Courses=C_max$node_list$label[idx],
                        BLocking_Factor = C_max$node_list$bf[idx],
                        Delay_Factor = C_max$node_list$df[idx],
                        Centrality = C_max$node_list$cf[idx],
                        Structural_Complexity = C_max$node_list$sc[idx])

  
  
  idx <- !(C_min$node_list$label %in% C_max$node_list$label)
  min_tab<- data.frame(Min_Only_Courses=C_min$node_list$label[idx],
                       BLocking_Factor = C_min$node_list$bf[idx],
                        Delay_Factor = C_min$node_list$df[idx],
                        Centrality = C_min$node_list$cf[idx],
                        Structural_Complexity = C_min$node_list$sc[idx])
  
    output$max_only_courses <- renderTable(kable(max_tab),
                                           striped = TRUE,
                                           align = "c")
    output$min_only_courses <- renderTable(kable(min_tab),
                                           striped = TRUE,
                                           align = "c")
}

shinyApp(ui, server)


```


